---
title: "One hot encoding"
author: "J(w)-cis-iun"
#date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(error=TRUE,echo=TRUE,message=FALSE,warning=FALSE)
library(reticulate)
library(tidyverse)
library(caret)
#reticulate::py_config()
```
```{python include=FALSE,eval=FALSE}
import pandas as pd
print(pd.__file__)
```
Most of the machine learning algorithms requires numerical features. One hot encoding turns each category in a categorical variable into a new feature, referred to as dummy variable. The values in each dummy variable are 0s and 1s. Each row only has one dummy variable being $1$ and the rest dummy variables are $0$. 

We can do one hot encoding manually or call the functions from **sklearn** or **caret**:

- Call **OneHotEncoder** from **sklearn.preprocessing**
- Call **dummyVars** from **caret** package

## One-hot in R

- By calling **dummyVars**, fit an object `dummifier` to categorical features in a data frame which need to be encoded;
- By calling **predict**, run `dummifier` with a formual on the data frame.
- The return of predict is a matrix, containing all the dummy variables.

1. Generate a categorical feature
```{r}
cats <- c("4G", "fiber", "phone", "landline")
N_samples <- 20
df <- tibble(technology = cats[sample(1:length(cats), N_samples, replace=TRUE)])
head(df$technology, 3)
```
2. Dummify the feature
```{r echo=TRUE}
dummifier <- dummyVars("~ technology", data=df)
dummy <- as_tibble(predict(dummifier, df))
knitr::kable(head(dummy, 3))
```

3. Append the dummy variables to the data frame
```{r echo=TRUE}
df <- df %>% add_column(dummy, .after ="technology") 
str(df)  
```

## One-hot in Python
```{python}
import numpy as np
import pandas as pd
from sklearn.preprocessing import OneHotEncoder
```
```{python}
N_samples = 20
cats = np.array(["4G", "fiber", "phone", "landline"])
inds = np.random.default_rng().integers(low=0, high=len(cats), size=N_samples)
df = pd.DataFrame(cats[inds], columns=['technology'])
df
```
```{python}
encoder = OneHotEncoder()
result = encoder.fit_transform(df[["technology"]]).toarray()
print(result)
print(encoder.get_feature_names(['technology']))
```
